Reasoning with Graphs for Quantum Computation

### 2008-03-07 papers/reasoning-graphically/paper.tex
typos
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1416 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

The representation of Compact Closed Categories as Graphs, discussed
in~\S\ref{sec:graph-repr-comp}, is too restrictive for reasoning about
quantum computation. In particular, graphical rules such the Spider
Theorem, from Figure~\ref{fig:spider}-left, are frequently needed, but
not expressible. In this section we combine the {\em !-boxes} and {\em
  variable-node} extensions to graphs. We call this representation
\emph{graph patterns}. This forms a representation that allows us to
express, in a finite way, the infinite family of concrete graphs
equations. In particular, the Spider Theorem can now be represented as
shown in Figure~\ref{fig:spider}-right. We also extend the notion of
matching for graph pattens.
%as well as present a corresponding notion of plugging. 
This provides the foundations for the rewriting machinery
in~\S\ref{sec:rewriting} which can then be used to reasoning about
quantum computation.

@@

Fortunately, it is quite easy to provide decidable matching: the size
of the unfolding that needs to be considered can be bounded. The key
observation is that a graph $G_1$ will never match a graph with less
non-variable nodes. Thus unfolding of $G_1$ can be bounded by the
number of non-variable nodes in $G_2$. While this gives a generate and
test style algorithm, it is not efficient. The intuition for an
efficient algorithm is to search through one graph incrementally
increasing the matched part and avoids symmetries in matching.  
%The development of an efficient algorithm is ongoing work.

>>>>

The representation of Compact Closed Categories as graphs, discussed
in~\S\ref{sec:graph-repr-comp}, is too restrictive for reasoning about
quantum computation. In particular, graphical rules such the spider
theorem, from Figure~\ref{fig:spider}-left, are frequently needed, but
not expressible. In this section we combine the {\em !-boxes} and {\em
  variable-node} extensions to graphs. We call this representation
\emph{graph patterns}. This forms a representation that allows us to
express, in a finite way, the infinite family of concrete graphs
equations. In particular, the Spider Theorem can now be represented as
shown in Figure~\ref{fig:spider}-right. We also extend the notion of
matching for graph pattens.
%as well as present a corresponding notion of plugging. 
This provides the foundations for the rewriting machinery
in~\S\ref{sec:rewriting} which can then be used to reasoning about
quantum computation.

@@

Fortunately, it is quite easy to provide decidable matching: the size
of the unfolding that needs to be considered can be bounded. The key
observation is that a graph $G_1$ will never match a graph with fewer
non-variable nodes. Thus unfolding of $G_1$ can be bounded by the
number of non-variable nodes in $G_2$. While this gives a generate and
test style algorithm, it is not efficient. The intuition for an
efficient algorithm is to search through one graph incrementally
increasing the matched part and avoids symmetries in matching.  
%The development of an efficient algorithm is ongoing work.
